Revealing transient strain in geodetic data with Gaussian process regression

### 2017-10-14 RevisedSubmission/hines_hetland_2017_gji.tex
did more

We assume a diffuse prior for the components of $\mitbf{m}_i$, that is to say $\mitbf{m}_i \sim \mathcal{N}(\mitbf{0},\kappa^2\mitbf{I})$ in the limit as $\kappa \to \infty$. Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$. However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior. Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study. We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally corelated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$. For example, $\mitbf{\eta}_i$ can be uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}. The appropriate noise model may vary depending on the application, and we hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$, until Section \ref{sec:NoiseModel}. We are now able to write the distributed of $\mitbf{d}_i$ as 
\begin{equation}\label{eq:DataDist}
\mitbf{d}_i \sim \mathcal{N}\left(\mitbf{0}, C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i}  + \kappa^2\mitbf{G}\mitbf{G}^T \right),
\end{equation}
where $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$.

We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$. That is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$. A general solution for $\hat{u}_i$ is derived in \citet[sec. 8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{GP}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function

@@

\begin{equation}\label{eq:PosteriorCov2}

@@

@@

It should be noted that we ignore any covariances between the easting and northing components, which reduces the computational complexity of our analysis because we can determine each component of the posterior independently. However, we are inherently assuming that the principle components of transient displacements and the noise processes are aligned with the cardinal directions, which is admittedly an arbitrary assumption.

>>>>

We assume a diffuse prior for the components of $\mitbf{m}_i$, that is to say $\mitbf{m}_i \sim \mathcal{N}(\mitbf{0},\kappa^2\mitbf{I})$ in the limit as $\kappa \to \infty$. Of course, the secular velocities, $b_i^{(j)}$, are spatially correlated and we could invoke a tectonic model to form a prior on $b_i^{(j)}$. However, in our application to the Pacific Northwest, we will be using displacement time series which are long enough to sufficiently constrain $b_i^{(j)}$ for each station, avoiding the need to incorporate a prior. Likewise, seasonal deformation is spatially correlated \citep{Dong2002,Langbein2008}, and it may be worth exploring and exploiting such a correlation in a future study. We assume that $\mitbf{\eta}_i$ is a spatially and/or temporally correlated random vector distributed as $\mathcal{N}(\mitbf{0},\mitbf{C}_{\eta_i})$. For example, $\mitbf{\eta}_i$ can be uncorrelated white noise, temporally correlated noise describing benchmark wobble \citep[e.g.,][]{Wyatt1982,Wyatt1989}, and/or spatially correlated noise describing common mode error \citep[e.g.,][]{Wdowinski1997}. The appropriate noise model may vary depending on the application, and we hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$, until Section \ref{sec:NoiseModel}. We are now able to write the distribution of $\mitbf{d}_i$ as 
\begin{equation}\label{eq:DataDist}
\mitbf{d}_i \sim \mathcal{N}\left(\mitbf{0}, C_{u_i}(\mitbf{P},\mitbf{P}) + \mitbf{C}_{\eta_i}  + \kappa^2\mitbf{G}\mitbf{G}^T \right),
\end{equation}
where $C_{u_i}(\mitbf{P},\mitbf{P})$ represents the matrix $[C_{u_i}(p,p')]_{(p,p') \in \mitbf{P} \times \mitbf{P}}$.

We form a posterior estimate of transient displacements, denoted as $\hat{u}_i$, by updating $u_i$ with the fact that $\mitbf{d}_i^*$ was realized from the random vector $\mitbf{d}_i$, that is to say $\hat{u}_i = u_i |(\mitbf{d}_i = \mitbf{d}^*_i)$. A general solution for $\hat{u}_i$ is derived in \citet[sec. 8.9]{VonMises1964}, where we find that $\hat{u}_i$ is distributed as $\mathcal{GP}(\mu_{\hat{u}_i},C_{\hat{u}_i})$ with mean function

@@

\begin{multline}\label{eq:PosteriorCov2}
C_{\hat{u}_i}(p,p') = C_{u_i}(p,p') - \\ 

@@

\end{multline}

@@

It should be noted that we have ignored any covariances between the easting and northing components of $\vec{u}$ and $\vec{\mitbf{d}}$. This simplification reduces the computational complexity of evaluating the posterior transient displacements because each component can be evaluated independently. However, we are inherently assuming that the principle axes describing the distribution of $\vec{u}(p)$ and $\vec{d}^{(jk)}$ are aligned with the cardinal directions, which is admittedly an arbitrary assumption.
